Data Specialization for a Staged Lambda Calculus

### 2015-02-22 paper/example.tex

\section {Example: Staged Quickselect in \texorpdfstring{\lang}{λ12}}
\label{sec:staging}

@@

\ur{
\begin{itemize}
\item 
I find it hard to distinguish between blue and black on printed paper.
\item
$\pause$ should be called ``promote'' or something like that. I
  believe this is what Taha calls cross stage persistence.  THat work should
  be cited.
\end{itemize}
}

@@

Suppose that we wish to issue order statistics queries on a collection of items
represented as a list \texttt{l}. We can use quickselect \cite{quickselect}, an
expected linear time algorithm which takes a list \texttt{l} and a rank
\texttt{k} and returns the element with rank \texttt{k} in \texttt{l}.

>>>>

\section {Example: Staged Quickselect in \texorpdfstring{\lang}{λ12}}
\label{sec:example}

@@

@@

Suppose that we wish to issue order statistics queries on a collection of items
represented as a list \texttt{l}. We can use quickselect \cite{Hoare:1961}, an
expected linear time algorithm which takes a list \texttt{l} and a rank
\texttt{k} and returns the element with rank \texttt{k} in \texttt{l}.

### 2015-02-22 paper/implementation.tex

We also have a number of optimizations which drastically improve the readability
of split programs. We split patterns directly, rather than translating them into
lower-level constructs. Finally, we perform many local simplifications during
the splitting algorithm, most notably, not pairing precomputations when one is
known to be $\tup{}$.

>>>>

In the splitting algorithm, we perform a number of optimizations which
drastically improve the readability of split programs. For example, we split
patterns directly, instead of first translating them into lower-level
constructs. We also take advantage of many local simplifications, most notably,
not pairing precomputations when one is known to be $\tup{}$.

### 2015-02-22 paper/intro.tex

\emph{Program specialization}, or partial evaluation, is a well-known
specialization technique which, given $a$, transforms $f$ into a new function
$f_a(-)$ which computes $f(a,-)$. This transformation essentially substitutes
$a$ for the first argument of $f$, then evaluates in place any subexpressions of
$f$ depending only on that argument.

\emph{Data specialization} \cite{knoblock96} is a technique for specializing $f$
\emph{without} the fixed argument $a$, instead splitting $f$ into a pair of
functions $f_1$ and $f_2$. $f_1(a)$ produces a data structure containing the
results of the computations which depend only on $a$; $f_2$ then completes the
computation, given this data structure and the varying argument; that is,
$f_2(f_1(a),-)$ computes $f(a,-)$.

>>>>

\emph{Program specialization}, or partial evaluation \cite{futamura71,jones96},
is a well-known specialization technique which, given $a$, transforms $f$ into a
new function $f_a(-)$ which computes $f(a,-)$. This transformation essentially
substitutes $a$ for the first argument of $f$, then evaluates in place any
subexpressions of $f$ depending only on that argument.

\emph{Data specialization} \cite{knoblock96,jorring86} 
is a technique for specializing $f$ \emph{without} the fixed argument $a$,
instead splitting $f$ into a pair of functions $f_1$ and $f_2$. $f_1(a)$
produces a data structure containing the results of the computations which
depend only on $a$; $f_2$ then completes the computation, given this data
structure and the varying argument; that is, $f_2(f_1(a),-)$ computes $f(a,-)$.
